---
title: "Trust in PBS"
author: "Christopher Ali, Hilde Van den Bulck, Jonathan Kropko"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: journal
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Abstract

This contribution explores the how and why of trust in PBS. After this
introduction, we sketch the constitutive characteristics of PBS. Next,
we theoretically explore the notion of trust in general and with regards
to media in particular, resulting in a set of more specific research
questions. Subsequently, we discuss the methodological set up. We
explored US audiences' trust in PBS by means of a survey questionnaire.
Next, we discuss the results, focusing on institutional aspects, on
characteristics of the offerings and performance of PBS and on the role
of long-standing relationships between PBS and its audiences. Finally,
we discuss the implications of our understanding of what constitutes
trust in PBS for the institution and for government policy towards PBS
in an age of distrust in media and society.

# Analyses {.tabset .tabset-fade}

## Initialization of the dataset


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(summarytools)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(ggwordcloud)
library(tidytext)
library(plotly)
```

```{r, message=FALSE, warning=FALSE}
pbs <- read_csv("data_clean.csv")
```

The following table provides summary statistics for all variables in the
data:

```{r, warning=FALSE}
print(dfSummary(pbs, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')
```

## Trust in PBS

For the following analyses we will use the following dependent variables:

* Trust in public institutions including PBS
* Value for taxpayers money
* Aspects of PBS that contribute to trust (arts&ed, public funding….)
* Variables detailing what PSB news, kids, arts and entertainment are considered good at
* Open questions

Independent variables can include

* age, 
* political orientation, 
* income, 
* race/ethnicity, 
* other media use

### How and in what ways is trust in PBS based in its institutional / structural characteristics (funding, independence)

The following figure reports aggregate levels of trust for each of 12 American institutions.
```{r, fig.width=12, fig.height=8}
pbs_trust <- select(pbs, starts_with("personaltrust")) %>%
  rename("NPR"="personaltrust_npr",
         "NASA"="personaltrust_nasa",                
         "Smithsonian"="personaltrust_smithsonian",
         "PBS"="personaltrust_pbs",
         "Public Schools"="personaltrust_publicschools",
         "Congress"="personaltrust_congress",
         "Current President"="personaltrust_currentpresident",
         "State Government"="personaltrust_stategovernment",
         "Postal Services"="personaltrust_postalservices",
         "Courts of Law"="personaltrust_courtsoflaw",         
         "Law Enforcement and Police"="personaltrust_lawenforcementpolice",
         "Military"="personaltrust_military") %>%
  pivot_longer(c("NPR","NASA","Smithsonian","PBS","Public Schools",
               "Congress","Current President","State Government",
               "Postal Services","Courts of Law","Law Enforcement and Police","Military"),
               names_to = "institution",
               values_to = "trust") %>%
  group_by(institution, trust) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(trust = fct_relevel(trust, "No opinion","Distrust a great deal",
                             "Somewhat distrust", "Neither trust nor distrust",
                             "Somewhat trust", "Trust a great deal"),
         institution = relevel(as.factor(institution),ref="PBS")) %>%
  na.omit()

g <- ggplot(pbs_trust, aes(y=trust, x=count, fill=institution)) +
  geom_col() +
  facet_wrap(~institution) + 
  guides(fill="none") +
  theme(text=element_text(family="serif")) +
  xlab("Count") +
  ylab("Level of trust")
g

```

We can also conduct a comparative analysis across institutions: which institutions are most similar in terms of how individuals rate them comparably?

```{r}
pbs_trust <- select(pbs, starts_with("personaltrust")) %>%
  rename("NPR"="personaltrust_npr",
         "NASA"="personaltrust_nasa",                
         "Smithsonian"="personaltrust_smithsonian",
         "PBS"="personaltrust_pbs",
         "Public Schools"="personaltrust_publicschools",
         "Congress"="personaltrust_congress",
         "Current President"="personaltrust_currentpresident",
         "State Government"="personaltrust_stategovernment",
         "Postal Services"="personaltrust_postalservices",
         "Courts of Law"="personaltrust_courtsoflaw",         
         "Law Enforcement and Police"="personaltrust_lawenforcementpolice",
         "Military"="personaltrust_military") 

newlevels <- c("1"="Distrust a great deal", "2"="Somewhat distrust",
               "3"="Neither trust nor distrust", "4"="Somewhat trust",
               "5"="Trust a great deal", "NA"="No opinion")

pbs_trust <- pbs_trust %>%
  mutate(NPR = fct_recode(NPR, !!!newlevels),
         NASA = fct_recode(NASA, !!!newlevels),
         Smithsonian = fct_recode(Smithsonian, !!!newlevels),
         PBS = fct_recode(PBS, !!!newlevels),
         `Public Schools` = fct_recode(`Public Schools`, !!!newlevels),
         Congress = fct_recode(Congress, !!!newlevels),
         `Current President` = fct_recode(`Current President`, !!!newlevels),
         `State Government` = fct_recode(`State Government`, !!!newlevels),
         `Postal Services` = fct_recode(`Postal Services`, !!!newlevels),
         `Courts of Law` = fct_recode(`Courts of Law`, !!!newlevels),
         `Law Enforcement and Police` = fct_recode(`Law Enforcement and Police`, !!!newlevels),
         Military = fct_recode(Military, !!!newlevels)) %>%
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.character, as.numeric)

distance <- dist(t(pbs_trust), method="euclidian")
```

```{r}
fit <- cmdscale(distance,eig=TRUE, k=2)
map <- as.data.frame(fit$points) %>%
  mutate(institution = rownames(fit$points))
map

trustmeans <- data.frame(institution = names(colMeans(pbs_trust, na.rm=TRUE)),
                         avg_trust = colMeans(pbs_trust, na.rm=TRUE))

map <- full_join(map, trustmeans, by="institution")
map
```

```{r}
g <- ggplot(map, aes(institution=institution, x=V1, y=V2, color=`avg_trust`)) +
  geom_point() +
  xlab("Dimension 1") +
  ylab("Dimension 2") +
  ggtitle("A map of institutions with regard to trust") +
  scale_color_gradient(low='red', high='blue')
ggplotly(g)
```

### How and in what ways is trust in PBS based on its programmatic and content-based decisions?

```{r, warning=FALSE, message=FALSE}
contrib <- select(pbs, starts_with("contrib")) %>%
  select(-contributetrustpbs_other, -contributetrustpbs_othertext)
contrib <- contrib %>%
  pivot_longer(cols = colnames(contrib),
               names_to = "reason",
               values_to = "trust") %>%
  group_by(reason, trust) %>%
  summarize(count = n()) %>%
  mutate(reason = fct_recode(reason,
                             "Adult Programs"="contributetrustpbs_adultprograms",
                             "Arts and Culture"="contributetrustpbs_artsandculture",
                             "Children's Programs"="contributetrustpbs_childrenprograms",
                             "Entertainment"="contributetrustpbs_entertainment",
                             "Funding"="contributetrustpbs_funding",
                             "Kids' Games"="contributetrustpbs_kidsgames",
                             "Local News"="contributetrustpbs_localnews",
                             "Local Stations"="contributetrustpbs_localstation",
                             "National News"="contributetrustpbs_nationalnews",
                             "Organization"="contributetrustpbs_organisation")) %>%
  na.omit() %>%
  mutate(trust = fct_relevel(trust, 
                             "No opinion", 
                             "Distrust a lot", 
                             "Somewhat distrust", 
                             "Neither trust nor distrust", 
                             "Somewhat trust",
                             "Trust a lot" ))

```

```{r, fig.width=12, fig.height=6}
g <- ggplot(contrib, aes(y=trust, x=count, fill=reason)) +
  geom_col() +
  facet_wrap(~ reason, ncol=5) +
  guides(fill=FALSE) +
  theme(text=element_text(family="serif")) +
  xlab("Count") +
  ylab("Level of trust")
g
```


### How and in what ways is trust in PBS  based on processed-based, historical viewing habits?



### How is trust in PBS related to political affiliations, to media use and attitudes and to wider SES?
 


## Role of PBS in trust in news in post-truth era

For the following analyses we will use the following dependent variables:

* Why do you trust PBS: position of national and local news in reasons to trust PBS
* Open ended questions: role of news in why people trust PBS
* How much do you trust PBS (news) - comparing with trust in other news outlets
* What characteristics of PBS news (thoroughness, distinctiveness..) contribute to that trust
 
Independent variables can include

* age, 
* political orientation, 
* income, 
* race/ethnicity, 
* other media use

### What is the position of news in the overall trust of PBS

### What is said about news in the open questions – news is biggest category

### How does trust in PBS news compare with trust of other mainstream news organizations

### What aspects of pbs news are most trusted by those who watch

### How do these things differ according to age, political orientation, income, race/ethnicity, other media use
 

 
