---
title: "Trust in PBS"
author: "Christopher Ali, Hilde Van den Bulck, Jonathan Kropko"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: journal
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This contribution explores the how and why of trust in PBS. After this
introduction, we sketch the constitutive characteristics of PBS. Next,
we theoretically explore the notion of trust in general and with regards
to media in particular, resulting in a set of more specific research
questions. Subsequently, we discuss the methodological set up. We
explored US audiences' trust in PBS by means of a survey questionnaire.
Next, we discuss the results, focusing on institutional aspects, on
characteristics of the offerings and performance of PBS and on the role
of long-standing relationships between PBS and its audiences. Finally,
we discuss the implications of our understanding of what constitutes
trust in PBS for the institution and for government policy towards PBS
in an age of distrust in media and society.

# Loading and summarizing the data

We load the following libraries:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(summarytools)
```

The cleaned survey data exist in the following CSV file:

```{r, message=FALSE, warning=FALSE}
pbs <- read_csv("data_clean.csv")
```

The following table provides sumamry statistics for all variables in the
data:

```{r, warning=FALSE}
print(dfSummary(pbs, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')
```

# Analyses

## Characteristics of trustee

Institution-based trust is impersonal and based on the conviction that
society provides structures (regulations, institutions) that allow
people to trust those structures and, through that, each other. As
McKnight & Chervany (2013) explain, this comes with a structural
assurance: structures have built-in "guarantees" that ensure trusting
intentions.

Characteristics of PBS as trusted institution:

-   R1: How and in what ways is trust in PBS based in its institutional
    / structural characteristics (funding, independence)

Comparing all the institutions in terms of level of trust:

```{r, fig.width=12, fig.height=8}
pbs_trust <- select(pbs, starts_with("personaltrust")) %>%
  rename("NPR"="personaltrust_npr",
         "NASA"="personaltrust_nasa",                
         "Smithsonian"="personaltrust_smithsonian",
         "PBS"="personaltrust_pbs",
         "Public Schools"="personaltrust_publicschools",
         "Congress"="personaltrust_congress",
         "Current President"="personaltrust_currentpresident",
         "State Government"="personaltrust_stategovernment",
         "Postal Services"="personaltrust_postalservices",
         "Courts of Law"="personaltrust_courtsoflaw",         
         "Law Enforcement and Police"="personaltrust_lawenforcementpolice",
         "Military"="personaltrust_military") %>%
  pivot_longer(c("NPR","NASA","Smithsonian","PBS","Public Schools",
               "Congress","Current President","State Government",
               "Postal Services","Courts of Law","Law Enforcement and Police","Military"),
               names_to = "institution",
               values_to = "trust") %>%
  group_by(institution, trust) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(trust = fct_relevel(trust, "No opinion","Distrust a great deal",
                             "Somewhat distrust", "Neither trust nor distrust",
                             "Somewhat trust", "Trust a great deal"),
         institution = relevel(as.factor(institution),ref="PBS")) %>%
  na.omit()

g <- ggplot(pbs_trust, aes(y=trust, x=count, fill=institution)) +
  geom_col() +
  facet_wrap(~institution) + 
  guides(fill="none") +
  theme(text=element_text(family="serif")) +
  xlab("Count") +
  ylab("Level of trust")
g

```

-   Process-based trust results from repeated contact and interaction
    between trustor and trustee. While the original contact and
    interaction may be coincidental (stumbling upon a particular media
    channel), repeated contacts can result in the reduction of
    uncertainty and the building of trust. o Translated into how long
    folks have been PBS viewers (Mr Rogers' effect) o R2: How and in
    what ways is trust in PBS based on processed-based, historical
    viewing habits?
-   Character-based trust, finally, results from perceived
    trustworthiness of particular characteristics of the trustee by the
    trustor. o Quality of programs o R3: How and in what ways is trust
    in PBS based on its programmatic and content-based decisions?
    Characteristics of the trustor
-   SES, political leanings, wider media use.. o R4: How is trust in PBS
    related to political affiliations, to media use and attitudes and to
    wider SES?

(Trust in) PBS news in a polarized news media ecology how does PBS news
perform on characteristics of 'good journalism' (trustworthiness,
balance...) compared to other news media how does this differ according
to characteristics of the audience (SES, political leanings, media use)
