---
title: "Trust in PBS"
author: "Christopher Ali, Hilde Van den Bulck, Jonathan Kropko"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: journal
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This contribution explores the how and why of trust in PBS. After this
introduction, we sketch the constitutive characteristics of PBS. Next,
we theoretically explore the notion of trust in general and with regards
to media in particular, resulting in a set of more specific research
questions. Subsequently, we discuss the methodological set up. We
explored US audiences' trust in PBS by means of a survey questionnaire.
Next, we discuss the results, focusing on institutional aspects, on
characteristics of the offerings and performance of PBS and on the role
of long-standing relationships between PBS and its audiences. Finally,
we discuss the implications of our understanding of what constitutes
trust in PBS for the institution and for government policy towards PBS
in an age of distrust in media and society.

# Loading and summarizing the data

We load the following libraries:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(summarytools)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
```

The cleaned survey data exist in the following CSV file:

```{r, message=FALSE, warning=FALSE}
pbs <- read_csv("data_clean.csv")
```

The following table provides sumamry statistics for all variables in the
data:

```{r, warning=FALSE}
print(dfSummary(pbs, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
      method = 'render')
```

# Analyses

## Characteristics of trustee

Institution-based trust is impersonal and based on the conviction that
society provides structures (regulations, institutions) that allow
people to trust those structures and, through that, each other. As
McKnight & Chervany (2013) explain, this comes with a structural
assurance: structures have built-in "guarantees" that ensure trusting
intentions.

Characteristics of PBS as trusted institution:

-   R1: How and in what ways is trust in PBS based in its institutional
    / structural characteristics (funding, independence)

Comparing all the institutions in terms of level of trust:

```{r, fig.width=12, fig.height=8}
pbs_trust <- select(pbs, starts_with("personaltrust")) %>%
  rename("NPR"="personaltrust_npr",
         "NASA"="personaltrust_nasa",                
         "Smithsonian"="personaltrust_smithsonian",
         "PBS"="personaltrust_pbs",
         "Public Schools"="personaltrust_publicschools",
         "Congress"="personaltrust_congress",
         "Current President"="personaltrust_currentpresident",
         "State Government"="personaltrust_stategovernment",
         "Postal Services"="personaltrust_postalservices",
         "Courts of Law"="personaltrust_courtsoflaw",         
         "Law Enforcement and Police"="personaltrust_lawenforcementpolice",
         "Military"="personaltrust_military") %>%
  pivot_longer(c("NPR","NASA","Smithsonian","PBS","Public Schools",
               "Congress","Current President","State Government",
               "Postal Services","Courts of Law","Law Enforcement and Police","Military"),
               names_to = "institution",
               values_to = "trust") %>%
  group_by(institution, trust) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  mutate(trust = fct_relevel(trust, "No opinion","Distrust a great deal",
                             "Somewhat distrust", "Neither trust nor distrust",
                             "Somewhat trust", "Trust a great deal"),
         institution = relevel(as.factor(institution),ref="PBS")) %>%
  na.omit()

g <- ggplot(pbs_trust, aes(y=trust, x=count, fill=institution)) +
  geom_col() +
  facet_wrap(~institution) + 
  guides(fill="none") +
  theme(text=element_text(family="serif")) +
  xlab("Count") +
  ylab("Level of trust")
g

```

* Process-based trust results from repeated contact and interaction between trustor and trustee. While the original contact and interaction may be coincidental (stumbling upon a particular media channel), repeated contacts can result in the reduction of uncertainty and the building of trust. Translated into how long folks have been PBS viewers (Mr Rogers' effect) 

```{r}
pbs_whytrust <- select(pbs, open_whyyoutrust, personaltrust_pbs) %>%
  mutate(most_trust = (personaltrust_pbs == 'Trust a great deal')) %>%
  group_by(open_whyyoutrust) %>%
  summarize(frequency = n(),
            most_trust = round(100*mean(most_trust), 1)) %>%
  arrange(-frequency)

knitr::kable(pbs_whytrust)
```
```{r, fig.width=10, fig.height=10, cache=TRUE}
library(ggwordcloud)
library(tidytext)
data(stop_words)
trustwords <- read_csv("PBS categories - open_whyyoutrust(1).csv") 
tw_word <- trustwords %>%
  mutate(`New category` = fct_recode(`New category`, "News exhibits political moderation, \n unbiasedness, or accuracy" = "News exhibits political moderation, unbiasedness, or accuracy")) %>%
  unnest_tokens(word, `Old category`) %>%
  anti_join(stop_words) %>%
  group_by(`New category`) %>%
  mutate(catcount = n()) %>%
  ungroup() %>%
  group_by(`New category`, word) %>%
  summarize(count=n(),
            catcount = mean(catcount),
            count = count/catcount)

g <- ggplot(tw_word, aes(label = word, size=count, color=count)) +
  geom_text_wordcloud_area() +
  facet_wrap(~ `New category`, scales="free") +
  scale_size_area(max_size = 20) 
g
```

Paper 1: Trust in PBS
RQ
- How and in what ways is trust in PBS based in its institutional / structural characteristics (funding, independence)
- How and in what ways is trust in PBS  based on processed-based, historical viewing habits?
- How and in what ways is trust in PBS based on its programmatic and content-based decisions?
- How is trust in PBS related to political affiliations, to media use and attitudes and to wider SES?
 
Dependent variables
- Trust in public institutions including PBS
- Value for taxpayers money
- Aspects of PBS that contribute to trust (arts&ed, public funding….)
- Variables detailing what PSB news, kids, arts and entertainment are considered good at
- Open questions
Independent variables
- age, political orientation, income, race/ethnicity, other media use

Paper 2: Role of PBS in trust in news in post-truth era
RQ
- What is the position of news in the overall trust of PBS
- What is said about news in the open questions – news is biggest category
- How does trust in PBS news compare with trust of other mainstream news organizations
- What aspects of pbs news are most trusted by those who watch
- How do these things differ according to age, political orientation, income, race/ethnicity, other media use
 
Dependent variables:
- Why do you trust PBS: position of national and local news in reasons to trust PBS
- Open ended questions: role of news in why people trust PBS
- How much do you trust PBS (news) - comparing with trust in other news outlets
- What characteristics of PBS news (thoroughness, distinctiveness..) contribute to that trust
 
Independent variables
- age, political orientation, income, race/ethnicity, other media use
 
